\documentclass[11pt, class=article, crop=false]{standalone}
\usepackage[subpreambles=true]{standalone}
\usepackage[T1]{fontenc} % for font setting
\usepackage{newtxtext,newtxmath}
\usepackage{import,
            graphicx,
            parskip,
            url,
            amsmath,
            wrapfig,
            fancyhdr,
            soul,
            tabularx,
            authblk,
            textcomp,
            lineno,
            hyperref}


% for special characters in bibliography            
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
       
% margin
\usepackage[top=2.54cm, bottom=2.54cm, left=2.54cm, right=2.54cm]{geometry}

\linenumbers
\date{} 
\title{}

\begin{document}

\section*{Significance} % for oikos

Our manuscript capitalizes on deepening our understanding of movement by integrating advanced modeling (theory) with behavioral ecology (empirical). This work combines multiple facets of science, continuing to close the knowledge gap between why species move and whether it is a non-random behavior. Our results stand out among others due to the use of multiple types of explanatory movement drivers (extrinsic, intrinsic, and environmental) across multiple species from multiple families all within one model structure. For these reasons, we feel \textit{Oikos} is the best outlet for publication. 

\section*{Data Archiving}

All data and R code can be found on the GitHub repository under the following link: \url{https://github.com/ashleylaroque/public_cmr_movement}

\section*{Ethics}

The fieldwork conducted for this study follows IACUC protocol for vertebrate animals (ID: 2022-1173).

\section*{Acknowledgements}

This work is based on materials supported by University of North Carolina at Greensboro Startup, the Faculty Internal Award, John O'Brien Ecological Field Award, and Dorothy Levis Munroe Research Award funds. We graciously thank our undergraduate students for their help in conducting seasonal fieldwork and digitizing data. We are grateful to University of North Carolina at Greensboro and North Carolina Agricultural and Technical State University for their corroboration in granting us access to Gateway Research Park. 

\section*{Conflicts of Interest}

None declared.

\section*{Author Contributions}

A.T. and A.L. conceived the idea; A.L., S.K., and A.T. performed  field work; A.L. and A.T. performed statistical analysis and drafted the manuscript; All authors contributed to revision and gave final approval for  publication.



\end{document}